---
title: "Data Analytics Homework"
subtitle: "Exploring the FINRA Financial Capabilities Data Set"
author: 
  - "ECO 230: Business and Economics Research and Communication"
  - "Instructor: James M. Murray, Ph.D."

output: 
#  html_document:
#    keep_md: yes
  pdf_document: 
    fig_width: 7
    fig_height: 3.5
    
params:
  output_format: 'all'
urlcolor: blue
---

## Preparing the Data Set and the R Environment
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following command downloads loads a sample of the National Financial Capability Study 2015 survey responses. The data has been cleaned, coded, and stored in an R data frame called `df`.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
load(url("http://murraylax.org/datasets/findata.RData"))
```

See the code book for all the variable descriptions.

The questions below involve using the packages `tidyverse`, `Hmisc`, and `scales`. Install and load these into memory using, 
```{r, eval=FALSE, echo=TRUE}
install.packages("tidyverse")
install.packages("Hmisc")
```
```{r, message=FALSE, warning=FALSE}
library("tidyverse")
library("Hmisc")
library("scales")
```

## Questions

1. What is the scale of measurement and class for the education variable, `Edu`. Give the definition for this scale of measurement. If the scale of measurement is nominal or ordinal, use R to show the possible categories. Hint: See the statistics tutorial, "Introduction to Data".

2. Compute the mean number of correct quiz questions (variable `QuizScore`) for each educational level (variable `Edu`). Hint: See "Group and Pipe" from the "Introduction to Data" statistics tutorial.

3. Create a bar plot showing the mean rating (scale 1-7) that people give themselves on their financial knowledge (variable `FinancialKnow`) for each level of education (variable `Edu`). Give the plot error bars to display confidence intervals for each mean. Give the plot a title, make the bars dark red in color, and make sure labels do not overlap.

4. Create again a plot like #3, but make a multiple bar chart that includes separate bars for males and females (variable `Female` is equal to 1 for a female and is 0 otherwise). Hint: use `fill=as.factor(Female)` in the aesthetics layer.

When you take a mean of a binary variable (a variable equal to 0 or 1), the result is the proportion equal to one. Then you can convert this to a percentage. For example, the following R code computes the mean of `RetireContribute` which is a binary variable equal to 1 if a person regularly makes contributions to his or her retirement savings:

```{r}
mean(df$RetireContribute, na.rm=TRUE)
```

A mean equal to 0.67 means that 67\% of people contribute to their retirement savings.

5. Create a bar plot of the proportion of people that contribute to their retirement (i.e. use `stat_summary()` with `fun.data=mean_sdl`), for people with each possible quiz score (0-6) (use `as.factor(QuizScore)` for the x aesthetic). Have your plot include the following:

 - Dark green bars
 
 - Make sure the scale for the y-axis is a percentage. Use `scale_y_continuous(label=percent)`.
 
 - Give the plot a title
 
 - Give error bars to visualize the confidence intervals for these estimates
 
 - Finally, give a one or two sentence description on what information is revealed in the plot about the relationship between financial literacy and making retirement contributions.
 
 
